---
title: "Visual story telling part 2: flights at ABIA"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE,warning=FALSE,include=FALSE}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(xts)
library(fpp)
library(ggmap)
library(dplyr)
library(plyr)
library(reshape2)
library(egg)
options(warn=-1)
```

```{r,echo=FALSE}
Airport=read.csv("~/Desktop/ABIA.csv")
```

```{r,include=FALSE}
#Removing the outliers 
Airport=Airport[-c(60414,98309),]
```

```{r,include=FALSE}
AUSdeps=Airport[Airport$Origin=="AUS",]
AUSdeps_counts=count(AUSdeps,'Dest')
AUSdeps2 <- AUSdeps_counts[order(-AUSdeps_counts$freq),] 
AUSdepsplot=AUSdeps2[AUSdeps2$freq>700,]
```

```{r,include=FALSE}
AUSdepsplot$Dest <- factor(AUSdepsplot$Dest, levels = AUSdepsplot$Dest[order(-AUSdepsplot$freq)])

Top_Destinations=ggplot(AUSdepsplot,aes(Dest,y=freq))+
  geom_col(na.rm=TRUE,color="red",fill="blue")+
  labs(x="Destinations",y="Total flights in 2008",title="Top Destinations from Austin in 2008")+
  theme(panel.background = element_blank(),
                  panel.grid.minor = element_blank(),
                  axis.ticks  = element_blank(),
                  axis.line   = element_line(colour=NA),
                  axis.line.x = element_line(colour="grey80")) 
```
```{r,echo=FALSE}
AUSdepsDAL=AUSdeps[AUSdeps$Dest=="DAL",]
AUSdepsDFW=AUSdeps[AUSdeps$Dest=="DFW",]
AUSdepsIAH=AUSdeps[AUSdeps$Dest=="IAH",]
AUSdepsPHX=AUSdeps[AUSdeps$Dest=="PHX",]
AUSdepsDEN=AUSdeps[AUSdeps$Dest=="DEN",]
TopAUSdeps=rbind(AUSdepsDAL,AUSdepsDFW,AUSdepsIAH,AUSdepsPHX,AUSdepsDEN,deparse.level = 1)
```

```{r,echo=FALSE}
Topdep=ggplot(TopAUSdeps,aes(y=ArrDelay,x=factor(DayOfWeek),fill=factor(DayOfWeek)))+
  geom_bar(stat="identity",position = "dodge",na.rm=TRUE)+
  facet_grid(factor(TopAUSdeps$Dest) ~ .)+
  coord_flip()+
  labs(y="Arrival Delay (minutes)",x="Day of the Week",title="Arrival Delay per Day of the Week",subtitle="Data for Top 5 destinations from Austin")

ArrDelPerDay = Topdep + scale_fill_manual(breaks=c("1", "2", "3","4","5","6","7"),values=c("grey46", "red", "grey46", "grey46", "grey46", "grey46", "grey46"), labels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"),name="Day Of the Week")
```

### We are going to try to find the day with the most arrival delays for flights from Austin Airport. First we will find the Top 5 Destinations from Austin for 2008. 
```{r,echo=FALSE}
Reasons=TopAUSdeps[TopAUSdeps$DayOfWeek==2,]
Reasons=Reasons[complete.cases(Reasons),]
DF=Reasons[,c(15,18,25:29)]
DF2=gather(DF,event,total,CarrierDelay:LateAircraftDelay)

Causes=ggplot(DF2,aes(Dest,total,fill=event))+
  geom_bar(stat="identity",position="fill",na.rm=TRUE)+
  theme_classic()+
  labs(title="Causes of Arrival Delays on Tuesdays",subtitle="Data for Top 5 Destinations from Austin",x="Destination",y="Total Delay Causes (Percentages)")+
  scale_fill_manual(values=c("red", "green", "blue", "yellow", "brown"),name="Cause of Delay")
```
```{r,echo=FALSE}
Top_Destinations
```








### For each of these 5 destinations we will find the arrival delay per day for each different destination. 
```{r,echo=FALSE}
ArrDelPerDay
```











### The majority of delays are on Tuesday for 3/5 destinations. Since this is a surprising insight lets find out why Tuesdays have the most delays for all 5 of the destinations. 
```{r,echo=FALSE}
Causes
```







### Most delays at these airports are because of late aircraft. If you are flying from Austin to DAL, DFW, or IAH (all in Texas) don't fly on a Tuesday if you don't want to risk being late. 


